---
title: "Basic experiments workflow for simple epidemiological models"
author: Anton Antonov
date: 2020-04-13
output: html_notebook
---

# Introduction

The primary purpose of this notebook is to give a “stencil workflows” for simulations using the package ["ECMMon-R](https://github.com/antononcube/ECMMon-R), [AAr2].

We present two code workflows: one based on "low-level" package usage, the other based on (monadic) pipelines.

The model in this notebook -- SEI2R -- differs from [the classical SEIR model](https://en.wikipedia.org/wiki/Compartmental_models_in_epidemiology) with the following elements:

1. Two separate infected populations: one is "severely symptomatic", the other is "normally symptomatic"

2. The monetary equivalent of lost productivity due to infected or died people is tracked

**Remark:** We consider the contagious disease propagation models as instances of the more general [System Dynamics (SD)](https://en.wikipedia.org/wiki/System_dynamics) models.
We use SD terminology in this notebook.

**Remark:** The SEI2R model is a modification of the classic epidemic model SEIR, [Wk1, HH1].

**Remark:** The monadic pipelines in the notebook can be used for attempts to calibrate SEI2R with real data. 
(For example, data for the [2019–20 coronavirus outbreak](https://en.wikipedia.org/wiki/2019â20_coronavirus_outbreak), [WRI1].)

**Remark:** The repository "ECMMon-R", [AAr2], has a [flexdashboard](https://github.com/antononcube/ECMMon-R/blob/master/flexdashboards/SEI2HR-flexdashboard.Rmd) 
for interactive experimentation with the package models. The flexdashboard can be easily developed further or changed in order to fit user's needs.

# Workflow

1. Get one of the classical epidemiology models.

2. Extend the model equations if needed or desired.

3. Set relevant initial conditions for the populations.

4. Pick model parameters to adjust and “play with.”

5. Using monadic pipelines (or an interactive interface) experiment with different values of the parameters.

    - In order to form “qualitative understanding.”

6. Get real life data.

    - Say, for the 2019-20 coronavirus outbreak.

7. Attempt manual or automatic calibration of the model.

    - This step will most likely require additional data transformations and programming.

    - Only manual calibration is shown in this notebook.

# Install and load `ECMMon` 

This notebook use the package `ECMMon-R`, [AAr2], which implements a software monad for rapid specification of Epidemiology Compartmental Modeling (ECM) workflows.

Monad's name is "ECMMon", which stands for "**E**pidemiology **C**ompartmental **M**odeling **Mon**ad".

```{r, message=FALSE}
devtools::install_github( "antononcube/ECMMon-R")
library(ECMMon)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Low-level code

## Getting the model code

Here we take the SEI2R model implemented in the package ["ECMMon-R"](https://github.com/antononcube/ECMMon-R), [AAr2]:

```{r}
model <- SEI2RModel( initialConditionsQ = TRUE, rateRulesQ = TRUE, birthsTermQ = FALSE )
```

The model object is a list with named elements:

```{r}
names(model)
```

Here are the stocks:

```{r}
as.data.frame(model$Stocks)
```

Here are the rates:

```{r}
as.data.frame(model$Rates)
```

Here is the Right Hand Side (RHS) function used in calls of the function `ode` of the package "deSolve", [KS1]:

```{r}
model$RHSFunction
```

Here are the initial conditions:

```{r}
as.data.frame(model$InitialConditions)
```

Here are the rate values:

```{r}
as.data.frame(model$RateRules)
```

## Parameter changes 

Here we change the initial condition for the stock Infected Severely Symptomatic Population (ISSP):

```{r}
model$InitialConditions["ISSPt"] <- 20
```


Here we change the contact rate for to be ISSP to be 0.5:

```{r}
model$RateRules["contactRateISSP"] <- 0.3
```

## Simulation

In this section we compute a simulation with model object described above. 
(Using [`deSolve::ode`](https://www.rdocumentation.org/packages/deSolve/versions/1.28/topics/ode).)

Here are the simulation times:

```{r}
times <- seq( 0, 365, 1)
```

Here we compute the simulation:

```{r}
sol <- deSolve::ode(y = model[["InitialConditions"]], times = times, func = model[["RHSFunction"]], parms = model[["RateRules"]], method = "rk4" )
```

Here we convert the matrix `sol` into a data frame:

```{r}
dfSol <- as.data.frame(sol)
colnames(dfSol) <- gsub( "time", "Time", colnames(dfSol) )
head(dfSol)
```

**Remark:** We can see in the first row that ISSP starts with the initial condition we set.

## Visualization

Next we plot the solutions of the "non-money" stocks using the package "ggplot2" after we first convert the data frame `dfSol` into long form using the package "tydir":

```{r}
dfSol %>% 
  tidyr::pivot_longer( cols = colnames(dfSol)[-1], names_to = "Stock", values_to = "Value" ) %>%
  dplyr::filter( Stock != "MLPt" ) %>% 
  ggplot2::ggplot( ) +
  ggplot2::geom_line( ggplot2::aes( x = Time, y = Value, color = Stock ) )
```

Here we plot the solutions histograms:

```{r}
hist(sol)
```

# Pipelines

Here is a pipeline that corresponds to the code in the previous section:

```{r}
ecmObj <- 
  ECMMonUnit( SEI2RModel() ) %>% 
  ECMMonAssignInitialConditions( initConds = c( "ISSPt" = 20 ) ) %>%
  ECMMonAssignRateValues(rateValues = c("contactRateISSP" = 0.3 ) ) %>% 
  ECMMonSimulate( 365 ) %>% 
  ECMMonPlotSolutions( stocksSpec = ".*Population" )
```

```{r}
ecmObj <- 
  ecmObj %>%
  ECMMonPlotSolutionHistograms(bins=20)
```

Here we see the first rows of the solutions data frame (in the monad object):

```{r}
head(ecmObj %>% ECMMonTakeSolution)
```

# Calibration with real data

It is important to calibrate these kind of models with real data, or at least to give a serious attempt to such a calibration. 
If the calibration is “too hard” or “impossible” that would indicate that the model is not that adequate. (If adequate at all.)

The calibration efforts can be (semi-)automated using special model-to-data goodness of fit functions and a minimization procedure. 
(See for example, [AA2].)

In this section we just attempt to calibrate SEI2R over real data taken from [WRI1] using a specialized interactive interface.

*(To be finished...)*

# References

## Articles

[Wk1] Wikipedia entry, ["Compartmental models in epidemiology"](https://en.wikipedia.org/wiki/Compartmental_models_in_epidemiology).

[HH1] Herbert W. Hethcote (2000). ["The Mathematics of Infectious Diseases"](http://leonidzhukov.net/hse/2014/socialnetworks/papers/2000SiamRev.pdf). SIAM Review. 42 (4): 599–653. Bibcode:2000SIAMR..42..599H. doi:10.1137/s0036144500371907.

[AA1] Anton Antonov, ["Coronavirus propagation modeling considerations"](https://github.com/antononcube/SystemModeling/blob/master/Projects/Coronavirus-propagation-dynamics/Documents/Coronavirus-propagation-modeling-considerations.md), (2020), [SystemModeling at GitHub](https://github.com/antononcube/SystemModeling).

[AA2] Anton Antonov, [Answer of "Model calibration with phase space data"](https://mathematica.stackexchange.com/a/190295/34008), (2019), [Mathematica StackExchage](https://mathematica.stackexchange.com).

## Repositories & packages

[WRI1] Wolfram Research, Inc., ["Epidemic Data for Novel Coronavirus COVID-19"](https://www.wolframcloud.com/obj/resourcesystem/published/DataRepository/resources/Epidemic-Data-for-Novel-Coronavirus-COVID-19), [WolframCloud](https://www.wolframcloud.com).

[KS1] Karline Soetaert et al, ["deSolve: Solvers for Initial Value Problems of Differential Equations ('ODE', 'DAE', 'DDE')"](https://cran.r-project.org/web/packages/deSolve/index.html), [CRAN](https://cran.r-project.org).

[AAr1] Anton Antonov, [Coronavirus propagation dynamics project](https://github.com/antononcube/SystemModeling/tree/master/Projects/Coronavirus-propagation-dynamics), (2020), [SystemModeling at GitHub](https://github.com/antononcube/SystemModeling).

[AAr2] Anton Antonov, [Epidemiology Compartmental Modeling Monad R package](https://github.com/antononcube/ECMMon-R), (2020), [ECMMon-R at GitHub](https://github.com/antononcube/ECMMon-R).

[AAp1] Anton Antonov, ["Epidemiology models Mathematica package"](https://github.com/antononcube/SystemModeling/blob/master/Projects/Coronavirus-propagation-dynamics/WL/EpidemiologyModels.m), (2020), [SystemModeling at GitHub](https://github.com/antononcube/SystemModeling).

[AAp2] Anton Antonov, ["Epidemiology models modifications Mathematica package"](https://github.com/antononcube/SystemModeling/blob/master/Projects/Coronavirus-propagation-dynamics/WL/EpidemiologyModelModifications.m), (2020), [SystemModeling at GitHub](https://github.com/antononcube/SystemModeling).

[AAp3] Anton Antonov, ["System dynamics interactive interfaces functions Mathematica package"](https://github.com/antononcube/SystemModeling/blob/master/WL/SystemDynamicsInteractiveInterfacesFunctions.m), (2020), [SystemModeling at GitHub](https://github.com/antononcube/SystemModeling).
