---
title: "Batch simulation example"
author: Anton Antonov
date: 2020-05-24
output: html_notebook
---

```{r}
library(ECMMon)
library(magrittr)
library(ggplot2)
```

# Introduction

Illustrate the batch simulation abilities of ECMMon.

# Basic pipeline

Here is the (single-site) model and a base simulation pipeline with it:

```{r}
ecmObj <- 
  ECMMonUnit( SEI2HRModel(initialConditionsQ = TRUE, rateRulesQ = TRUE) ) %>% 
  ECMMonSimulate( maxTime = 365 ) %>% 
  ECMMonPlotSolutions( stocksSpec = ".Population" )
```

# Batch simulation

## Different rates

See the model rates:

```{r}
ecmObj <- 
  ecmObj %>% 
  ECMMonEchoModelTableForm( part = "RateRules" )
```

Batch simulation:


```{r}
dfParams <- do.call( expand.grid, list( "aip" = seq(20, 40, 5), "aincp" = seq(6, 18, 6), "lpcr" = 0 ) )
dfParams
```


```{r}
system.time(
  dfRes <- 
    ecmObj %>% 
    ECMMonBatchSimulate( params = list( "aip" = seq(20, 40, 10), "aincp" = seq(6, 18, 6), "lpcr" = 0 ), maxTime = 365, resultForm = NULL )%>% 
    ECMMonTakeValue
)
```

```{r}
dfRes
```

Plot results:

```{r}
ggplot( dfRes ) +
  geom_line( aes( x = Time, y = Value, color = Stock ) ) +
  facet_wrap( ~ aip + aincp, ncol = 3 )
```

Sensitivity analysis: 

```{r}
ggplot( dfRes %>% dplyr::filter( Stock == "ISSPt" ) ) +
  geom_line( aes( x = Time, y = Value, color = factor(aip) ) ) +
  facet_wrap( ~ aincp, ncol = 1 ) 
```

## Different initial conditions

Here are the model stocks:

```{r}
(ecmObj %>% ECMMonTakeSingleSiteModel)$Stocks
```

```{r}
ecmObj <- 
  ecmObj %>% 
  ECMMonEchoModelTableForm( "Stocks" )
```

Here are the initial conditions:

```{r}
ecmObj <- 
  ecmObj %>% 
  ECMMonEchoModelTableForm( "InitialConditions" )
```

Here is the parameters data frame for different initial conditions:

```{r}
dfParameters <- do.call( expand.grid, list( "TP0" = seq(2, 3, 1) * 10^5, "ISSPt" = 10 ^ seq(0, 2, 1) ) )
dfParameters <- cbind( dfParameters, "SPt" = dfParameters$TP0 - dfParameters$ISSPt,  "lpcr" = 0  )
dfParameters
```

Note that we added the column "SPt" in order to have consistent initial conditions.

Batch simulation:

```{r}
dfRes <- 
  ecmObj %>% 
  ECMMonBatchSimulate( params = dfParameters, maxTime = 365 )%>% 
  ECMMonTakeValue
```

```{r}
dfRes
```

Plot results:


```{r}
ggplot( dfRes ) +
  geom_line( aes( x = Time, y = Value, color = Stock ) ) +
  facet_wrap( ~ TP0 + ISSPt, labeller = "label_both", ncol = 3, scales = "free" )
```

Sensitivity analysis:

```{r}
ggplot( dfRes %>% dplyr::filter( Stock == "ISSPt" ) ) +
  geom_line( aes( x = Time, y = Value, color = interaction( TP0, ISSPt ) ) )
```

# Using parametric functions (first)

***This approach does not work.***

```{r, eval=FALSE }
ql <- 21
lsFuncs <- 
  purrr::map( seq(10, 120, 5), function(qsd) { 
    function(t) { 0.56 * ifelse( qsd <= t && t <= qsd + ql, 0.25, 1 ) } 
  })
```


```{r, eval=FALSE}
ecmObj2 <- 
  ECMMonUnit( SEI2HRModel( initialConditionsQ = TRUE, rateRulesQ = TRUE ) ) %>% 
  ECMMonBatchSimulate( params = list( "contactRateINSP" = lsFuncs ) , maxTime = 365 )%>% 
  ECMMonTakeValue
```


```{r, eval=FALSE}
ecmObj2 <- 
  ECMMonUnit( SEI2HRModel( initialConditionsQ = TRUE, rateRulesQ = TRUE ) ) %>% 
  ECMMonEcho("Simulation starts...") %>% 
  ECMMonAssignRateValues( 
    c( 
      "contactRateINSP" =  function(t) { 0.56 * ifelse( qsd <= t && t <= qsd + ql, 0.25, 1 ) },
      "contactRateISSP" =  function(t) { 0.56 * ifelse( qsd <= t && t <= qsd + ql, 0.25, 1 ) }
    ) 
  ) %>% 
  ECMMonAddRates( rateDescriptions = c( "qsd" = "Quaranine start day", "ql" = "Quarantine length"), rateValues = c( "qsd" = 100, "ql" = 14 ) ) %>% 
  ECMMonSimulate( 365, method = "rk4" ) %>% 
  ECMMonTakeDeSolveSolution
  ECMMonBatchSimulate( params = list( "contactRateINSP" = lsFuncs ) , maxTime = 365 )%>% 
  ECMMonTakeValue
```
