---
title: "Batch simulation examples with ERTMon"
author: Anton Antonov
date: 2020-05-25
output: html_notebook
---

```{r setup}
library(ECMMon)
library(ERTMon)
library(magrittr)
library(ggplot2)
```

# Introduction

In this notebook we illustrate the batch simulation abilities of ECMMon.

Here are the steps:

1. Formulate a basic workflow pipeline

2. Do batch simulations 

  - The batch simulations results are exported to data structures used by Event Records Transformations Monad (ERTMon).

3. Using the ERTMon object visualize batch simulations results.

# Basic pipeline

Here is the (single-site) model and a base simulation pipeline with it:

```{r}
ecmObj <- 
  ECMMonUnit( SEI2HRModel(initialConditionsQ = TRUE, rateRulesQ = TRUE) ) %>% 
  ECMMonSimulate( maxTime = 365 ) %>% 
  ECMMonPlotSolutions( stocksSpec = ".Population" )
```


# Batch similations

Here is a data frame for the batch simulations parameters:

```{r}
dfParameters <- do.call( expand.grid, list( "TP0" = seq(2, 3, 1) * 10^5, "ISSPt" = 10 ^ seq(0, 2, 1) ) )
dfParameters <- cbind( dfParameters, "SPt" = dfParameters$TP0 - dfParameters$ISSPt,  "lpcr" = 0  )
dfParameters
```

Batch simulate and export the data in "star schema" data structures that can be used by ERTMon:

```{r}
lsRes <- 
  ecmObj %>% 
  ECMMonBatchSimulate( params = dfParameters, maxTime = 365, resultForm = "StarSchema" )  %>% 
  ECMMonTakeValue
```

```{r}
names(lsRes)
```

```{r}
lsRes[[1]]
```

# Visualize batch simulations results

```{r}
lsVars <- unique(lsRes$EventRecords$Variable) 
dfCompSpec <- ERTMonEmptyComputationSpecification( nrow = length(lsVars) )
dfCompSpec$Variable <- lsVars
dfCompSpec$Aggregation.interval.length <- 1
dfCompSpec$Max.history.length <- 365
dfCompSpec$Normalization.function <- "None"
dfCompSpec$Normalization.scope <- "None"
dfCompSpec
```

```{r}
ertObj <- 
  ERTMonUnit( eventRecords = lsRes$EventRecords, entityAttributes = lsRes$EntityAttributes, compSpec = dfCompSpec ) %>% 
  ERTMonProcessEventRecords( alignmentSpec = "MinTime" )
```


```{r}
ertObj2 <- 
  ertObj %>% 
  ERTMonPlotFeatureMatrices( matrixNames = c( "ISSPt.Mean", "INSPt.Mean"), facets = "EntityID", scales = "free" )
```

