---
title: "Multi-site SEI2R over a grid graph"
author: Anton Antonov
date: 2020-05-19
output: html_notebook
---

```{r}
library(ECMMon)
library(tidyverse)
library(igraph)
#devtools::install_github('thomasp85/gganimate')
library(gganimate)
```

# Introduction

# Main workflow

## Make the graph

In this section we do the model extension and simulation over a regular grid graph with a constant traveling patterns matrix.

Here we create a grid graph with directed edges:

```{r}
m = 7; n = 12;
grGrid <- igraph::make_lattice( c(m,n), directed = TRUE )
igraph::layout_on_grid(grGrid)
plot(grGrid)
```


Note that:

- There is one directed edge between any two edge-connected nodes

- All horizontal edges point in one direction

- All vertical edges point in one direction

- The edges are directed from nodes with smaller indexes to nodes with larger indexes.

Here we make a constant traveling matrix and summarize it:


```{r}
matGridTravel <- as_adjacency_matrix(grGrid, sparse = TRUE) # or just grGrid[]
matGridTravel <- matGridTravel * 1000
Matrix::image(matGridTravel)
```

```{r}
summary(matGridTravel@x)
```

## Scaling the model

Here we scale the SEI2R model with the grid graph constant traveling matrix:

```{r}
model1 <- SEI2RModel(initialConditionsQ = TRUE, rateRulesQ = TRUE)
```

```{r}
model1 <- ECMMonUnit(model1) %>% ECMMonAssignRateValues( rateValues = c( "contactRateISSP" = 0.5, "contactRateINSP" = 0.5 ) ) %>% ECMMonTakeSingleSiteModel
```

```{r}
modelGrid <- ToSiteCompartmentalModel( model = model1, mat = matGridTravel, migratingStocks = ("SPt") ) 
```


```{r}
system.time(
  ecmObj <- 
    ECMMonUnit(model1) %>% 
    ECMMonAssignRateValues( rateValues = c("lpcr" = 0, "TP0" = 10^6, "contactRateISSP" = 0.5, "contactRateINSP" = 0.5, "aip" = 26, "aincp" = 6 ) ) %>% 
    ECMMonAssignInitialConditions( initConds = c("ISSPt" = 0, "INSPt" = 0, "SPt" = 10^6-1 ) ) %>% 
    ECMMonExtendByAdjacencyMatrix( mat = matGridTravel, migratingStocks = c( "SPt", "EPt", "RPt" ) ) %>% 
    ECMMonAssignInitialConditions( initConds = c("ISSP_1_t" = 1, "INSP_1_t" = 0 ) ) %>% 
    ECMMonSimulate( maxTime = 365, method = "rk4" )
)
```

```{r}
ecmObj <- 
  ecmObj %>% 
  ECMMonPlotSolutions( stocksSpec = "_1_", stockDescriptionsQ = FALSE, separatePlotsQ = FALSE  )
```

```{r}
ecmObj <- 
  ecmObj %>% 
  ECMMonPlotSolutions( stocksSpec = paste0( "_", m*n, "_"), stockDescriptionsQ = FALSE, separatePlotsQ = FALSE  )
```

```{r}
dfSolutions <- 
  ecmObj %>% 
  ECMMonGetSolutionLongForm( siteIdentifiersQ = TRUE ) %>% 
  ECMMonTakeValue
dfSolutions
```

```{r}
dfPlotQuery <- 
  dfSolutions %>% 
  dplyr::filter( Description == "Infected Normally Symptomatic Population") %>% 
  dplyr::mutate( nSiteID = as.numeric(SiteID) ) %>% 
  dplyr::mutate( i = (nSiteID - 1) %/% m + 1, j = nSiteID - m * ( nSiteID - 1 ) %/% m ) %>% 
  dplyr::group_by( SiteID ) %>% 
  dplyr::mutate( ValueNormalized = Value / max(Value) ) %>% 
  dplyr::ungroup() %>% 
  dplyr::arrange( Time, nSiteID, Stock )
dfPlotQuery
```

```{r}
ggplot( dfPlotQuery %>% dplyr::filter( Time %/% 5 == Time / 5 ) ) +
  geom_tile( aes( x = i, y = j, fill = ValueNormalized ) ) +
  scale_fill_distiller(palette = "Spectral") +
  facet_wrap( ~Time )
```


```{r}
ggaPlot <- 
  ggplot( dfPlotQuery %>% dplyr::filter( Time %/% 5 == Time / 5 ) ) +
  geom_tile( aes( x = i, y = j, fill = ValueNormalized ) ) +
  scale_fill_distiller(palette = "Spectral") +
  gganimate::transition_time( time = Time )
```

```{r}
gganimate::animate(ggaPlot, renderer = ffmpeg_renderer(), width = 480, height = round( 480 * m / n ) )
```

# Grid graph values

```{r}
# Set up resolution and palette.
my_resolution <- 100
my_palette    <-  colorRampPalette(c('blue','red'))

# This gives you the colors you want for every point.
my_max    <-  15
my_vector <-  1:15 / my_max
my_colors <-  my_palette(my_resolution)[as.numeric(cut(my_vector, breaks=my_resolution))]
```

```{r}
plot.igraph(grGrid, vertex.color = my_colors  )
```

